Robust Adaptation to Multi-Scale Climate Variability

### 2019-03-28 report-earths-future/figs/flowchart.tex
Update presentation text
put stuff into separate supp info file using latest terrible template from agu

\begin{tikzpicture}[node distance = 0.7cm, auto, font=\sffamily]
  % Place nodes
  \node [parameter] (genmod) {Generating Model};
  \node [parameter, above=of genmod] (N) {$N$};
  \node [parameter, below=of genmod] (M) {$M$};

@@

  \node [result, right=of historical] (pThat) {Estimated Exceedances per Year};
  \node [result, right= of future] (pT) {Observed Exceedances per Year};

  \node [parameter, right=of pThat] (fitting) {Estimating Model};
  \node [parameter, right=of seq] (threshold) {Flood Design Level};

@@

>>>>

\begin{tikzpicture}[node distance = 0.7cm, auto, font=\sffamily]
  % Place nodes
  \node [parameter] (genmod) {Scenario};
  \node [parameter, above=of genmod] (N) {$N$};
  \node [parameter, below=of genmod] (M) {$M$};

@@

  \node [calculated, right=of historical] (pThat) {Estimated Exceedances per Year};
  \node [calculated, right= of future] (pT) {Observed Exceedances per Year};

  \node [parameter, above=of pThat] (fitting) {Estimating Model};
  \node [parameter, right=of seq] (threshold) {Flood Design Level};

  \node [result, right=of pThat] (bias) {bias of mean};
  \node [result, right=of pT] (variance) {bias of variance};

@@

  \path[arrow] (pThat) -- (bias);
  \path[arrow] (pT) -- (bias);
  \path[arrow] (pThat) -- (variance);
  \path[arrow] (pT) -- (variance);

### 2019-03-28 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
put stuff into separate supp info file using latest terrible template from agu

\documentclass[
  draft,
  linenumbers
]{agujournal2018}
\journalname{Earth's Future}

% Macros and newcommands
\usepackage{xspace}
\newcommand{\eg}{e.g.\@\xspace}
\newcommand{\ie}{i.e.\@\xspace}
\newcommand{\normal}{\mathcal{N}}
\makeatletter
\newcommand*{\etc}{%
    \@ifnextchar{.}%
        {etc}%
        {etc.\@\xspace}%
}
\makeatother

% Package calls with options
\usepackage[letterpaper, margin=1in]{geometry} % page geometry
\usepackage[american]{babel} % language setting
\usepackage[inline]{enumitem} % in-line lists

% Other package calls
\usepackage{
  amssymb, amsmath,   % for math symbols
  apacite,            % AGU style citations apparently
  booktabs,           % for nice tables
  graphicx,           % for figures
  physics,            % for physics notation
  siunitx,            % for SI notation
  trackchanges,       % to track any new text
}

@@

This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, \eg,][]{Merz:2014gf}.
\add{A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures} \citep{lempert:2007,Borgomeo:2018hk} \add{and which fail along non-catastrophic modes} \citep{brown:2010}.
Although climate risk has been traditionally managed with centrally planned structural instruments (\eg, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
\remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (\eg, a stormwater barrier)}, operational\add{ (\eg, improved evacuation routes)}, and financial\add{ (\eg, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
These instruments are not typically implemented in isolation or statically.
Instead, investment decisions made at each point in time affect the viability, costs, and benefits of future decisions, causing the system to trace a ``pathway'' through time \citep{Walker:2013gi,Haasnoot:2013im,Haasnoot:2015dh}.

@@

Typical climate risk management policies do not use a single risk mitigation instrument, but rather build a portfolio of several instruments.
Each has its own operational period, which may or may not match the planning horizon of the portfolio as a whole.
This means that even if the portfolio has a long planning period, \ie if long-term plans are a priority, this goal may be best accomplished through a series of flexible and adaptive instruments with short individual planning periods.
\add{For example, the action that New York City can take today which best protects against uncertain hurricane risks over the 21st century could potentially be to purchase insurance and defer more permanent allocation of capital until some uncertainties are resolved.}
The costs and benefits of each individual instrument will be assessed over its individual, finite planning period, but decisions about the portfolio structure are evaluated over the longer planning horizon.

@@

Climate risk is governed by a variety of physical processes which occur on scales ranging from local and transient to global and permanent.
Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on \add{some} river floods, droughts, hurricanes, urban flooding, and many other climate hazards \remove{are}\add{has been the subject of substantial investigation} \remove{well documented} \citep[\eg,][]{Coumou:2012bc,Milly:2008dg,OGorman:2009hj,Trenberth:2003bj}.
Human activities can also affect climate risk through modification of local land or river systems \citep[see][]{Merz:2014gf}, and through changes in exposure to extremes \citep{baldassarre:2018,Jongman:2012cr}.
In combination, these effects highlight that the past may not be an adequate representation of future climate risk \citep[termed ``nonstationarity'' by][]{Milly:2008dg}.

\remove{Permanent or secular change is not, however, a prerequisite for the inadequacy of relying on historical records.}
\add{Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.}
\annote{One well-studied example}{This whole paragraph is added, but the track changes packages doens't allow math or references.} in the hydrological literature is the Hurst phenomenon, $\frac{R_n}{S_n} \sim n^h$, where $h \in [0.5, 1]$, $R_n$ is the range of the first $n$ data points, and $S_n$ is the sample standard deviation \citep{Hurst:1951tc,oconnell:2016}.
A variety of theories, including infinite memory, nonstationarity of the process, and specific storage mechanisms have been put forth to explain the Hurst phenomenon, none satisfactorily \citep{klemes:1974}.
Links between local hydrological processes and large-scale climate patterns, often referred to as teleconnections \citep{Angstrom:1935ej}, have also been used to explain \gls{lfv} in hydroclimate systems.
If these climate modes vary on decadal or longer time scales, such as the Pacific Decadal Oscillation \citep{Mantua:1997kj}, then this \gls{lfv} may appear in local hydrologgical records as well, leading the past to be a poor predictor of future behavior.
For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \acrfull{lfv}, due, \eg, to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
When records are short, and whether low-frequency external forcing or local memory dominates, it is easy to mistake the Hurst phenomenon for a secular trend which may not continue from the past into the future \citep{bhattacharya:1983}.

@@

The first step is to sample climate risk by generating synthetic streamflow sequences.
To do this, we model  annual-maximum flood peaks with a log-normal distribution, conditional on a location parameter which varies in time:
\begin{equation} \label{eq:log-normal}
  \log Q(t) \sim \normal \qty(\mu(t), \sigma(t)).
\end{equation}
We further assume a constant coefficient of variation of the log streamflow,
\begin{equation}
  \sigma(t) = \xi \mu(t)  
\end{equation}
and apply a lower threshold on the standard deviation
\begin{equation}
  \sigma(t) \geq \sigma_\text{min} > 0.
\end{equation}
This formulation describes the \begin{enumerate*}[label= (\roman*)]
  \item \gls{lfv} only;
  \item \gls{lfv} plus secular change; and
  \item secular change only
\end{enumerate*}
scenarios in a single equation; the only component which needs to change is the dependence of $\mu(t)$ on time, which we parameterize as
\begin{equation}\label{eq:nino3}
  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0),
\end{equation}
where $x(t)$ represents a climate time series which itself exhibits \gls{lfv} but not secular change.
This parameterization is analagous to the ``climate-informed'' approach described in several studies for estimating climate risk \citep{Delgado:2014ey,Merz:2014gf,Farnham:2018gs}.
Following \cref{eq:nino3}, when $\beta \neq 0$ there will be \gls{lfv}, and when $\gamma \neq 0$ there will be secular change.
The values of all parameters used for sampling climate risk are listed in \cref{sec:supp-methods} for each of the three scenarios considered.

We represent the climate state variable $x(t)$ through an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} as well as a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
We model \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
To create an annual time series, we average the October-December values of the NINO3 index for each year.
\Cref{sec:supp-nino-spectrum} shows a wavelet spectrum and time series plot of the resulting annual time series.
In \cref{sec:markov-generating} we consider an alternative parameterization of $\mu(t)$, \ie an alternative to \cref{eq:nino3}, which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.

@@

\remove{We finally consider}\add{Finally, we explicitly model} \remove{a HMM which explicitly represents}\gls{lfv} \add{using a} \gls{hmm}.
A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (\ie hidden) states $S(t)$ \citep{Rabiner:1986jk}.
The (unobserved) states evolve following a first-order Markov process, and the observed variable (\eg streamflow) depends only on the underlying state.
\Glspl{hmm} have been widely used for modeling streamflow sequences \citep{Bracken:2016ba} and \gls{enso} \citep{rojohernandez:2017}.
We fit  streamflow sequences using a \gls{hmm} with two states.
The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution \remove{conditional only on the state}\add{that is conditional only on the unobserved state variables}.
This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
For simplicity, we fit only a two-state \gls{hmm} to each sequence.
Future floods are then estimated by simulating future states from the estimated transition matrix and then drawing $Q(t)$ conditional on the simulated state.

@@

The log-normal trend model tends to over-estimate risk (positive bias), except when $N$ is large, because the model gives a non-zero probability to the trend being larger than it actually is.
The variance of these estimates is also large.
This again highlights the difficulty of fitting complex models for estimating risk when informational uncertainty is large.
By contrast, the stationary log-normal model and \gls{hmm}, which do not account for secular change, show relatively low variance of their estimates and exhibit low bias for short $M$.
As $N \rightarrow \infty$, these (mis-specified) models can only represent the trend by setting the scale parameter very large, leading to high estimation variance and (as $M \rightarrow \infty$) also a large bias.
This principle has prompted some to consider only the most recent years of the data, deliberately shortening $N$ \citep[\ie,][]{Muller:2014fc}.
However, these results also highlight that the increase in variance as $N$ is reduced may quickly outpace the utility of any bias reductions.

@@

Consistent with the conceptual illustration of \cref{fig:conceptual-sketch}, the results of \cref{fig:lfv-secular-nino3-bias-variance} highlight that the relative importance of secular change and \gls{lfv} depends on $M$.
When $M$ is long, climate risk is dominated by secular change and it becomes essential to model this risk explicitly with a more complex model (\ie, the linear trend model).
Alternatively, when $M$ is short, \gls{lfv} dominates and the increased variance associated with estimating a trend is not worth the modest reduction in bias.
As before, when the informational uncertainty is large (small $N$), the identifiability and predictability of the trend are  limited.
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{lfv-secular-nino3-bias-variance.pdf}
  \caption{
    As \cref{fig:secular-nino3-bias-variance} but for sequences generated with both \gls{lfv} and secular change.
  }\label{fig:lfv-secular-nino3-bias-variance}
\end{figure}

@@

For projects with longer $M$, \add{our results highlight the importance of identifying and predicting} \remove{it becomes necessary to consider the role of} secular change.
As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high intrinsic uncertainty in projections of future climate risk.
As the physical mechanisms cascade from global (\eg, global mean surface temperature) to regional  \citep[\eg, storm track position][]{Barnes:2015gl} and local (\eg, annual-maximum streamflows) scales, informational uncertainties also compound and increase \citep{Dittes:2017he}.
With sufficient information (large $N$), this informational uncertainty may be reduced, but this data cannot address intrinsic uncertainty and this zone is thus named the ``intrinsic uncertainty zone''.
Finally, if $N$ is limited then there will be strong potential for misleading estimates and over-extrapolation in this ``danger zone''.

@@

% -----------------------------------------------------------------------------
% SUPPLEMENTAL INFORMATION -- CURRENTLY INCLUDED IN THIS FILE FOR CONVENIENCE
% -----------------------------------------------------------------------------

\clearpage
\appendix

\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\theequation}{S\arabic{equation}}
\setcounter{equation}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

\section{Supplemental Methods}\label{sec:supp-methods}

In this section we provide further equations and parameter values used to generate and fit synthetic streamflow sequences.

\subsection{Sampling Climate Risk}\label{sec:supp-nino-spectrum}

As described in \cref{sec:methods-generating}, we use a synthetic \gls{enso} time series, specifically a NINO3 time series, from \citep{Ramesh:2016hf} to represent low-frequency climate variability.
\Cref{fig:enso-ts} shows the wavelet spectrum of this time series, calculated using the WaveletComp package \citep{Roesch:wlBQQoIs}.
The global (average) wavelet power spectrum shows a clear peak around 3-7 years, in line with previous studies of \gls{enso} although interactions with other modes of \acrfull{lfv} found in some studies \citep{Jin:1994wq} is dampened in this model.
There are, however, modes of \gls{lfv} between \SIrange{16}{128}{year} which are \emph{locally} strong at some points in time, which captures at least some effects of very low-frequency behavior for the purposes of this study.
\begin{figure}
  \includegraphics[width=\textwidth]{enso_wavelet.png}
  \caption{
    Wavelet analysis of the synthetic annual NINO3 time series described in \cref{sec:methods-generating}.
    (L): wavelet power spectrum.
    (R): global (average) power spectrum.
    Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
  }\label{fig:enso-ts}
\end{figure}

\subsection{Projecting Climate Risk over the Future $M$ Years}

As described in \cref{sec:methods}, we implement the stationary and trend log-normal models in a Bayesian framework using the stan programming language \citep{Carpenter:2017ke}.
The full model, including priors, for the stationary log-normal model is given by \cref{eq:log-normal-stationary}:
\begin{align}\label{eq:log-normal-stationary}
  \begin{split}
    \log Q_\text{hist} & \sim \normal \qty(\mu, \ \sigma) \\
    \mu &\sim \normal \qty(7, 1.5) \\
    \sigma &\sim \normal^+ \qty(1, 1)
  \end{split}
\end{align}
where $\normal$ denotes the normal distribution and $\normal^+$ denotes a half-normal distribution.
The full model, including priors, for the trend log-normal model is given by \cref{eq:log-normal-trend}:
\begin{align}\label{eq:log-normal-trend}
  \begin{split}
    \mu &= \mu_0 + \beta_\mu \qty(t - t_0) \\
    \log Q_\text{hist} & \sim \normal \qty(\mu, \ \xi \mu) \\
    \mu_0 & \sim \normal \qty(7, 1.5) \\
    \beta_\mu & \sim \normal \qty(0, 0.1) \\
    \log \xi & \sim \normal \qty(0.1, 0.1)
  \end{split}
\end{align}
where $\xi$ is an estimated coefficient of variation.

\subsection{Experiment Design}\label{sec:methods-experiments}

In this section we describe the parameters used to generate the specific sets of sequences that are analyzed in the results section.
To explore the temporal structure of risk, we generate $J=1000$ sequences of streamflow, for each of several combinations of $M$ and $N$.
Streamflow sequences are generated using the stationary generating functions both with and without superimposing a secular trend.

\begin{table}[ht]
  \centering
  \begin{tabular}{llll}
    \toprule
    Parameter & \gls{lfv} Only & Secular Only & Secular + \gls{lfv} \\
    \midrule
    $\mu_0$             & 6     & 6.5   & 6 \\
    $\gamma$            & 0     & 0.015 & 0.015\\
    $\beta$             & 0.5   & 0     & 0.5\\
    $\xi$               & 0.1   & 0.1   & 0.1\\
    $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
    Threshold           & 3000  & 3000  & 3000 \\
    \bottomrule
  \end{tabular}
  \caption{
    Parameters of \cref{eq:nino3} for the \gls{lfv} only, secular only, and secular plus \gls{lfv} scenarios.
  }\label{tab:nino-stationary}
\end{table}

\section{Sensitivity to climate generating mechanism}\label{sec:markov-generating}

In the main text we sampled climate risk by considering dependence on a single climate index, as shown in \cref{eq:nino3}.
Here we examine the sensitivity of our results to this assumption by considering a different parameterization for $\mu(t)$, specifically a two-state Markov chain model.
A Markov chain explicitly models transition between a fixed number of regimes, mimicking similar phenomena observed in nature.
The transition matrix is given by
\begin{equation}
  P = \mqty[\pi_1 & 1-\pi_1 \\ 1-\pi_2 & \pi_2].
\end{equation}
The transition matrix is first used to generate a sequence of states $S(t)$.
The value $\mu(t)$ depends only on $S(t)$ and on time itself:
\begin{equation}
  \mu(t) = \begin{cases}
    \mu_{1} + \gamma_1 \qty(t - t_0) & \qqtext{if} S(t) = 1 \\
    \mu_{2} + \gamma_2 \qty(t - t_0) & \qqtext{if} S(t) = 2
  \end{cases}
\end{equation}
For simplicity, we assume that the coefficient of variation is the same for both states and that $\pi_1=\pi_2$.
We further impose $\mu_{1} > \mu_{2}$ so that state 1 can be interpreted as the ``wet'' state and state 2 as the ``dry'' state.
Parameters for the three experiments are given in \cref{tab:markov-stationary}.
Results for these experiments are shown in \cref{fig:lfv-markov-bias-variance,fig:lfv-secular-markov-bias-variance,fig:secular-only-markov-bias-variance}.
\begin{table}[ht]
  \centering
  \begin{tabular}{llll}
    \toprule
    Parameter & \gls{lfv} Only & Secular Only & Secular + \gls{lfv} \\
    \midrule
    $\mu_1$             & 6.75  & 6.5   & 6.75\\
    $\mu_2$             & 6     & 6.5   & 6\\
    $\gamma_1$          & 0     & 0.015 & 0.015\\
    $\gamma_2$          & 0     & 0.015 & 0\\
    $\xi$               & 0.1   & 0.1   & 0.1\\
    $\sigma_\text{min}$ & 0.01  & 0.01  & 0.01\\
    Threshold           & 3000  & 3000  & 3000 \\
    \bottomrule
  \end{tabular}
  \caption{
    Parameters of the two-state Markov chain model for the \gls{lfv} only, secular only, and secular plus \gls{lfv} experiments.
  }\label{tab:markov-stationary}
\end{table}
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{lfv-only-markov-bias-variance.pdf}
  \caption{
    \Gls{lfv} only: as \cref{fig:lfv-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
  }\label{fig:lfv-markov-bias-variance}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{secular-only-markov-bias-variance.pdf}
  \caption{
    Secular change: as \cref{fig:secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
  }\label{fig:secular-only-markov-bias-variance}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{lfv-secular-markov-bias-variance.pdf}
  \caption{
    \Gls{lfv} plus secular change: as \cref{fig:lfv-secular-nino3-bias-variance} for sequences generated with the two-state Markov chain model.
  }\label{fig:lfv-secular-markov-bias-variance}
\end{figure}

>>>>

\documentclass[
  draft,
  linenumbers
]{agujournal2019}
\journalname{Earth's Future}

% Package calls with options
\usepackage[inline]{enumitem} % in-line lists

% Other package calls
\usepackage{
  amssymb, amsmath,   % for math symbols
  apacite,            % AGU style citations apparently
  booktabs,           % for nice tables
  graphicx,           % for figures
  natbib,             % for references
  physics,            % for physics notation
  siunitx,            % for SI notation
  trackchanges,       % to track any new text
  xspace,             % for ie and eg
}

@@

This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, e.g.\@\xspace,][]{Merz:2014gf}.
\add{A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures} \citep{lempert:2007,Borgomeo:2018hk} \add{and which fail along non-catastrophic modes} \citep{brown:2010}.
Although climate risk has been traditionally managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
\remove{Instead}\add{Rather}, factors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, a stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
These instruments are not typically implemented in isolation or statically.
Instead, investment decisions made at each point in time affect the viability, costs, and benefits of future decisions, causing the system to trace a ``pathway'' through time \citep{Walker:2013gi,Haasnoot:2013im,Haasnoot:2015dh}.

@@

Typical climate risk management policies do not use a single risk mitigation instrument, but rather build a portfolio of several instruments.
Each has its own operational period, which may or may not match the planning horizon of the portfolio as a whole.
This means that even if the portfolio has a long planning period, i.e.\@\xspace if long-term plans are a priority, this goal may be best accomplished through a series of flexible and adaptive instruments with short individual planning periods.
\add{For example, the action that New York City can take today which best protects against uncertain hurricane risks over the 21st century could potentially be to purchase insurance and defer more permanent allocation of capital until some uncertainties are resolved.}
The costs and benefits of each individual instrument will be assessed over its individual, finite planning period, but decisions about the portfolio structure are evaluated over the longer planning horizon.

@@

Climate risk is governed by a variety of physical processes which occur on scales ranging from local and transient to global and permanent.
Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on \add{some} river floods, droughts, hurricanes, urban flooding, and many other climate hazards \remove{are}\add{has been the subject of substantial investigation} \remove{well documented} \citep[e.g.\@\xspace,][]{Coumou:2012bc,Milly:2008dg,OGorman:2009hj,Trenberth:2003bj}.
Human activities can also affect climate risk through modification of local land or river systems \citep[see][]{Merz:2014gf}, and through changes in exposure to extremes \citep{baldassarre:2018,Jongman:2012cr}.
In combination, these effects highlight that the past may not be an adequate representation of future climate risk \citep[termed ``nonstationarity'' by][]{Milly:2008dg}.

Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
The Hurst phenomenon associated with stochastic long-memory processes has been invoked to explain long scale fluctuations in geophysics, physics, biology, medicine, traffic, network dynamics as well as for financial time series:
\begin{equation}
  \frac{R_n}{S_n} \sim n^h,
\end{equation}
where $h \in [0.5, 1]$, $R_n$ is the range of the first $n$ data points, and $S_n$ is the sample standard deviation \citep{Hurst:1951tc,oconnell:2016}.
A variety of theories, including infinite memory, nonstationarity of the process, and specific storage mechanisms have been put forth to explain the Hurst phenomenon, none fully satisfactorily \citep{klemes:1974}.
The Hurst exponent is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes, especially in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
\citet{lovejoy:2012,lovejoy:2013a} and \citet{lovejoy:2013} also argue that multifractal and long memory processes are an appropriate description of climate in general and weather extremes in particular. 
Based on this theory, regimes and quasi-periodic phenomena with increasing amplitude as their recurrence interval increases would be the norm rather than the exception.
Similar arguments are made by \citet{koutsoyiannis:2003}, \citet{markonis:2013}, and \citet{oconnell:2016}, and detailed discussions and some philosophical perspectives are offered by \citet{bloschl:2010} and \citet{montanari:2003}.
Wavelet methods have also been used to estimate the Hurst exponent \citep{simonsen:1998,chamoli:2007}, and to design simulation algorithms that reproduce self-similarity, long range dependence and quasi-periodic regimes \citep{Kwon:2007fj,bullmore:2001,geweke:1983,feng:2005}.

Links between local hydrological processes and large-scale climate patterns, often referred to as teleconnections \citep{Angstrom:1935ej}, have also been used to explain \gls{lfv} in hydroclimate systems.
If these climate modes vary on decadal or longer time scales, such as the Pacific Decadal Oscillation \citep{Mantua:1997kj}, then this \gls{lfv} may appear in local hydrologgical records as well, leading the past to be a poor predictor of future behavior.
For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \gls{lfv}, due, e.g.\@\xspace, to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
When records are short, and whether low-frequency external forcing or local memory dominates, it is easy to mistake the Hurst phenomenon for a secular trend which may not continue from the past into the future \citep{bhattacharya:1983}.

@@

The first step is to sample climate risk by generating synthetic streamflow sequences.
To do this, we model  annual-maximum flood peaks with a log-normal distribution, conditional on a location parameter which varies in time:
\begin{equation} \label{eq:log-normal}
  \log Q(t) \sim \mathcal{N} \qty(\mu(t), \sigma(t)).
\end{equation}
We further assume a constant coefficient of variation of the log streamflow,
\begin{equation}
  \sigma(t) = \xi \mu(t)  
\end{equation}
and apply a lower threshold on the standard deviation
\begin{equation}
  \sigma(t) \geq \sigma_\text{min} > 0.
\end{equation}
This formulation describes the \begin{enumerate*}[label= (\roman*)]
  \item \gls{lfv} only;
  \item \gls{lfv} plus secular change; and
  \item secular change only
\end{enumerate*}
scenarios in a single equation; the only component which needs to change is the dependence of $\mu(t)$ on time, which we parameterize as
\begin{equation}\label{eq:nino3}
  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0),
\end{equation}
where $x(t)$ represents a climate time series which itself exhibits \gls{lfv} but not secular change.
This parameterization is analagous to the ``climate-informed'' approach described in several studies for estimating climate risk \citep{Delgado:2014ey,Merz:2014gf,Farnham:2018gs}.
Following \cref{eq:nino3}, when $\beta \neq 0$ there will be \gls{lfv}, and when $\gamma \neq 0$ there will be secular change.
The values of all parameters used for sampling climate risk are listed in \cref{sec:supp-methods} for each of the three scenarios considered.

We represent the climate state variable $x(t)$ through an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} as well as a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
We model \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
To create an annual time series, we average the October-December values of the NINO3 index for each year.
\Cref{sec:supp-nino-spectrum} shows a wavelet spectrum and time series plot of the resulting annual time series.
In \cref{sec:markov-generating} we consider an alternative parameterization of $\mu(t)$, i.e.\@\xspace an alternative to \cref{eq:nino3}, which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.

@@

\remove{We finally consider}\add{Finally, we explicitly model} \remove{a HMM which explicitly represents}\gls{lfv} \add{using a} \gls{hmm}.
A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (i.e.\@\xspace hidden) states $S(t)$ \citep{Rabiner:1986jk}.
The (unobserved) states evolve following a first-order Markov process, and the observed variable (e.g.\@\xspace streamflow) depends only on the underlying state.
\Glspl{hmm} have been widely used for modeling streamflow sequences \citep{Bracken:2016ba} and \gls{enso} \citep{rojohernandez:2017}.
We fit  streamflow sequences using a \gls{hmm} with two states.
The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution \remove{conditional only on the state}\add{that is conditional only on the unobserved state variables}.
This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
For simplicity, we fit only a two-state \gls{hmm} to each sequence.
Future floods are then estimated by simulating future states from the estimated transition matrix and then drawing $Q(t)$ conditional on the simulated state.

@@

The log-normal trend model tends to over-estimate risk (positive bias), except when $N$ is large, because the model gives a non-zero probability to the trend being larger than it actually is.
The variance of these estimates is also large.
This again highlights the difficulty of fitting complex models for estimating risk when informational uncertainty is large.
By contrast, the stationary log-normal model and \gls{hmm}, which do not account for secular change, show relatively low variance of their estimates and exhibit low bias for short $M$.
As $N \rightarrow \infty$, these (mis-specified) models can only represent the trend by setting the scale parameter very large, leading to high estimation variance and (as $M \rightarrow \infty$) also a large bias.
This principle has prompted some to consider only the most recent years of the data, deliberately shortening $N$ \citep[i.e.\@\xspace,][]{Muller:2014fc}.
However, these results also highlight that the increase in variance as $N$ is reduced may quickly outpace the utility of any bias reductions.

@@

Consistent with the conceptual illustration of \cref{fig:conceptual-sketch}, the results of \cref{fig:lfv-secular-nino3-bias-variance} highlight that the relative importance of secular change and \gls{lfv} depends on $M$.
When $M$ is long, climate risk is dominated by secular change and it becomes essential to model this risk explicitly with a more complex model (i.e.\@\xspace, the linear trend model).
Alternatively, when $M$ is short, \gls{lfv} dominates and the increased variance associated with estimating a trend is not worth the modest reduction in bias.
As before, when the informational uncertainty is large (small $N$), the identifiability and predictability of the trend are  limited.
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{lfv-secular-nino3-bias-variance.pdf}
  \caption{
    As \cref{fig:secular-nino3-bias-variance} but for sequences generated with both \gls{lfv} and secular change.
  }\label{fig:lfv-secular-nino3-bias-variance}
\end{figure}

@@

For projects with longer $M$, \add{our results highlight the importance of identifying and predicting} \remove{it becomes necessary to consider the role of} secular change.
As illustrated schematically in \cref{fig:conceptual-sketch}, uncertainties as to future population, economic development, renewable energy technology, and greenhouse gas mitigation collectively lead to very high intrinsic uncertainty in projections of future climate risk.
As the physical mechanisms cascade from global (e.g.\@\xspace, global mean surface temperature) to regional  \citep[e.g.\@\xspace, storm track position][]{Barnes:2015gl} and local (e.g.\@\xspace, annual-maximum streamflows) scales, informational uncertainties also compound and increase \citep{Dittes:2017he}.
With sufficient information (large $N$), this informational uncertainty may be reduced, but this data cannot address intrinsic uncertainty and this zone is thus named the ``intrinsic uncertainty zone''.
Finally, if $N$ is limited then there will be strong potential for misleading estimates and over-extrapolation in this ``danger zone''.

@@
